Wireless smart monitoring energy system: a proof of concept

### 2021-03-15 paper/main.tex
server logic image

\subsection{Server}\label{SVR}
The server has been developed in Java 8 and managed using Maven. Tomcat is implemented to expose the servlet API. The server has the following responsibilities: to periodically broadcasts \textit{INIT messages} to let the plugs know its address; It listens for connections and updates from the plugs; it waits for acknowledgments from the plugs and, in case they are not received, it retransmits the lost packets; it exposes servlets to let the web interface get information about the plugs and send \textit{ON/OFF messages}. The periodic actions, namely, sending the \textit{INIT message} and retransmitting those that are lost, are implemented as extensions of \textit{Guava's AbstractScheduledService} which allows performing of these at constant intervals. Another thread waits and manages the plugs' responses. When a plug responds to the \textit{INIT message} the server will register its information and estimate the max power consumption for the appliance connected to the plug. This value is simply based on average values for the given appliance type as found on Google. When a plug sends an \textit{UPDATE message} the server will update its registered information, most important: the current energy consumption and the max energy consumption. Finally, the server waits for the acknowledgments of previously sent messages. These messages are kept with a timestamp until an acknowledgment with the same timestamp is received. If a configurable number of seconds elapses before this happens the packet will be sent again. The servlets simply allow the web interface to get the server's information about the plugs (the type of appliance connected, the current energy consumption, and more) and to turn ON or OFF any given plug. In this last case, a packet will be sent by the server and for it the server will wait for an acknowledgment. The information about the plugs saved by the server is kept as a \textit{thread-safe hashmap}. Some of this information, such as the IP address of the plugs, the type of appliance connected and the max power consumption is also saved in a \textit{XML file} to avoid loosing it in case of a server malfunction or restart.
\subsection{Web interface}\label{WI}
The web interface is written in React using Bootstrap and exposes a way to interact with the server to get 
the real-time power consumption for each plug and managing the connected devices. It performs some to HTTP, so over TCP, to the API exposed by the server through the servlet, performing polling calls. The GUI shown in Figure~\ref{fig:simulatedwebapp} presents a switch where the user can turn on or off the plug. If there is not enough energy available to turn on a device, the switch will be displayed as disabled in order to avoid a blackout.
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\linewidth]{assets/simulated_webapp}
	\caption{Web interface used to control the whole system}
	\label{fig:simulatedwebapp}
\end{figure}

\section{Methods}
Throughout the development, the communication between server and the plug has been simulated and tested using PacketSender which is a tool that allow to send a configurable packet through the network. To check the system's traffic and the correct handling of acknowledgments, we used PacketSender to simulate the traffic generated by the plug but not the acknowledgments and Wireshark to filter the traffic by port and type and check the correct retransmition of packets by the server. Finally the entire system as been tested to check the correct transmission of acknowledgments by the plugs.
\section{System results}
\section{Consumption}
%Qui mettiamo workflow generale con immagine algoritmo

>>>>

\subsection{Server}\label{SVR}
The server has been developed in Java 8 and managed using Maven. Tomcat is implemented to expose the servlet API. The server has the following responsibilities: to periodically broadcasts \textit{INIT messages} to let the plugs know its address; It listens for connections and updates from the plugs; it waits for acknowledgments from the plugs and, in case they are not received, it retransmits the lost packets; it exposes servlets to let the web interface get information about the plugs and send \textit{ON/OFF messages}. The periodic actions, namely, sending the \textit{INIT message} and retransmitting those that are lost, are implemented as extensions of \textit{Guava's AbstractScheduledService} which allows performing of these at constant intervals. Another thread waits and manages the plugs' responses. When a plug responds to the \textit{INIT message} the server will register its information and estimate the max power consumption for the appliance connected to the plug. This value is simply based on average values for the given appliance type as found on Google. When a plug sends an \textit{UPDATE message} the server will update its registered information, most important: the current energy consumption and the max energy consumption. Finally, the server waits for the acknowledgments of previously sent messages. These messages are kept with a timestamp until an acknowledgment with the same timestamp is received. If a configurable number of seconds elapses before this happens the packet will be sent again.
\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{assets/server_logic}
	\caption{Overview of server's logic for handling packets received by the plugs}
	\label{fig:server_logic}
\end{figure}
The servlets simply allow the web interface to get the server's information about the plugs (the type of appliance connected, the current energy consumption, and more) and to turn ON or OFF any given plug. In this last case, a packet will be sent by the server and for it the server will wait for an acknowledgment. The information about the plugs saved by the server is kept as a \textit{thread-safe hashmap}. Some of this information, such as the IP address of the plugs, the type of appliance connected and the max power consumption is also saved in a \textit{XML file} to avoid loosing it in case of a server malfunction or restart.

\subsection{Web interface}\label{WI}
The web interface is written in React using Bootstrap and exposes a way to interact with the server to get 
the real-time power consumption for each plug and managing the connected devices. It performs some to HTTP, so over TCP, to the API exposed by the server through the servlet, performing polling calls. The GUI shown in Figure~\ref{fig:simulatedwebapp} presents a switch where the user can turn on or off the plug. If there is not enough energy available to turn on a device, the switch will be displayed as disabled in order to avoid a blackout.
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\linewidth]{assets/simulated_webapp}
	\caption{Web interface used to control the whole system}
	\label{fig:simulatedwebapp}
\end{figure}

\section{Methods}
Throughout the development, the communication between server and the plug has been simulated and tested using PacketSender which is a tool that allow to send a configurable packet through the network. To check the system's traffic and the correct handling of acknowledgments, we used PacketSender to simulate the traffic generated by the plug but not the acknowledgments and Wireshark to filter the traffic by port and type and check the correct retransmition of packets by the server. Finally the entire system as been tested to check the correct transmission of acknowledgments by the plugs.

\section{System results}
\section{Consumption}
%Qui mettiamo workflow generale con immagine algoritmo
